---
title: "Rmarkdown Test on HUC data cleanup"
author: "Erik Beck, beck.erik@epa.gov"
date: "24 April 2025"
output: html_document
---
[comment]: # (Testing this comment)


# GRTS Info Processing After Retrieval #

Blah blah, say where in the processing chain this code set falls.


## Library Setup ##

```{r Library-Setup, message = FALSE, warning = FALSE }
library(skimr)

library(readxl)
library(dplyr)
# library(vioplot)
library(lubridate)
library(readr)
library(openxlsx)
library(feather)
```

## Path Setup ##


library(here)

STARTUP_DIR <- here()

setwd("..")

LOCAL_BASE_DIR <- getwd()
DATA_ROOT_DIR <- paste(LOCAL_BASE_DIR,'/DATA_IO/', sep="")

HUC_INPUT_DIR <- paste (DATA_ROOT_DIR,'HUC-Data-Lists/', sep="")
HUC_OUTPUT_DIR <- paste (DATA_ROOT_DIR,'HUC-Data-Lists/', sep="")

input_file <- paste(HUC_INPUT_DIR,'National_HUC12List.csv', sep="")



```{r file-path-setup}
infile <- '../DataOutput/GRTS-Data-NewEng-byHUC.pandas.xlsx'
# infile <- '../DataOutput/TESTGRTS-Data-NewEng-byHUC.pandas.xlsx' #test

R_outfile <- '../DataOutput/HUC-NewEng-cleaned.Rdata'
F_outfile <- '../DataOutput/HUC-NewEng-cleaned.feather'
Ex_outfile <- '../DataOutput/HUC-NewEng-cleaned.xlsx'

Ex_out_MA <- '../DataOutput/GRTS-MA.xlsx'
Ex_out_RI <- '../DataOutput/GRTS_RI.xlsx'
Ex_out_CT <- '../DataOutput/GRTS_CT.xlsx'
Ex_out_NH <- '../DataOutput/GRTS_NH.xlsx'
Ex_out_ME <- '../DataOutput/GRTS_ME.xlsx'
Ex_out_VT <- '../DataOutput/GRTS_VT.xlsx'

F_out_MA <- '../DataOutput/GRTS-MA.feather'
F_out_RI <- '../DataOutput/GRTS_RI.feather'
F_out_CT <- '../DataOutput/GRTS_CT.feather'
F_out_NH <- '../DataOutput/GRTS_NH.feather'
F_out_ME <- '../DataOutput/GRTS_ME.feather'
F_out_VT <- '../DataOutput/GRTS_VT.feather'


# comment missing
GRTS_df <- read_excel(path=infile)
```

## Data Type And Column Name Fixes ##

```{r Data-Type-And-Column-Name-Fixes}
# Name the first (left-most) column, which is a generated sequence number

colnames(GRTS_df)[1] <- 'data_seq'

# need to change these two date formats below to match upstream fixes
## probably ymd_hms()

print (GRTS_df$project_start_date)

GRTS_df$project_start_date <- ymd(GRTS_df$project_start_date)
GRTS_df$project_start_date <- mdy(GRTS_df$project_start_date_text)

GRTS_df$n_lbsyr <- as.numeric(GRTS_df$n_lbsyr)
GRTS_df$p_lbsyr <- as.numeric(GRTS_df$p_lbsyr)
GRTS_df$sed_tonsyr <- as.numeric(GRTS_df$sed_tonsyr)
GRTS_df$FakeHUC <- as.numeric(GRTS_df$huc_12)
GRTS_df$total_319_funds <- parse_number(GRTS_df$total_319_funds)
GRTS_df$project_dollars <- parse_number(GRTS_df$project_dollars)
GRTS_df$program_dollars <- parse_number(GRTS_df$program_dollars)
GRTS_df$epa_other <- parse_number(GRTS_df$epa_other)
GRTS_df$other_federal <- parse_number(GRTS_df$other_federal)
GRTS_df$state_funds <- parse_number(GRTS_df$state_funds)
GRTS_df$state_in_kind <- parse_number(GRTS_df$state_in_kind)
GRTS_df$local_funds <- parse_number(GRTS_df$local_funds)
GRTS_df$other_funds <- parse_number(GRTS_df$other_funds)
GRTS_df$local_in_kind <- parse_number(GRTS_df$local_in_kind)
GRTS_df$total_budget <- parse_number(GRTS_df$total_budget)
```


## Summary Stats ##
```{r Quick-Summary-Stats}
skim(GRTS_df)
```
## Write Data, All States Combined ##
```{r Write-Data}
save(GRTS_df, file=R_outfile)
write_feather(GRTS_df, F_outfile)
write.xlsx(GRTS_df,Ex_outfile)
```
## Split Data by State ##

```{r Hexfurcate}
GRTS_df %>% group_by(state) -> GRTS_byState
GRTS_byState %>% filter(state=="MA") -> GRTS_byState_MA
GRTS_byState %>% filter(state=="CT") -> GRTS_byState_CT
GRTS_byState %>% filter(state=="VT") -> GRTS_byState_VT
GRTS_byState %>% filter(state=="RI") -> GRTS_byState_RI
GRTS_byState %>% filter(state=="ME") -> GRTS_byState_ME
GRTS_byState %>% filter(state=="NH") -> GRTS_byState_NH
```

## Write Feather and Excel Data for Each State ##

### Excel ###
```{r State-Xlsx}
write.xlsx(GRTS_byState_MA, Ex_out_MA)
write.xlsx(GRTS_byState_ME, Ex_out_ME)
write.xlsx(GRTS_byState_CT, Ex_out_CT)
write.xlsx(GRTS_byState_VT, Ex_out_VT)
write.xlsx(GRTS_byState_RI, Ex_out_RI)
write.xlsx(GRTS_byState_NH, Ex_out_NH)
```

### Feather ###
```{r State-Feather}
write_feather(GRTS_byState_MA, F_out_MA)
write_feather(GRTS_byState_ME, F_out_ME)
write_feather(GRTS_byState_CT, F_out_CT)
write_feather(GRTS_byState_VT, F_out_VT)
write_feather(GRTS_byState_RI, F_out_RI)
write_feather(GRTS_byState_NH, F_out_NH)
```


## End ##

```{r Fini}

```


